---
title: "DBI compliance test results - comparative report"
author: "https://github.com/aryoda/R_DBI_compliance_reports"
date: '`r format(Sys.Date(), format = "%B %d, %Y")`'
output: 
  html_document: 
    number_sections: yes
    toc: yes
---

<style type="text/css">
div#TOC li {
    list-style:none;
    background-image:none;
    background-repeat:none;
    background-position:0; 
}
</style>


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(knitr)
require(kableExtra)
require(formattable)
require(ggplot2)
```



```{r ggplot themes, echo = FALSE}

# resusable theme for consistent ggplot layouts in this report
# (and easy to change - just here in one place)
report_theme <- function() {
  theme(plot.title = element_text(size = 10),
        axis.text = element_text(size = 8),
        strip.text = element_text(size = 8),
        axis.title = element_text(size = 8),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        legend.position = "bottom")
}
```



```{r ggplot_helper, echo = FALSE}

# ggplot label formatter function (see ?ggplot2::labellers and ?ggplot2::facet_grid for details)
# This function should inherit from the S3 class "labeller" using as_labeller()
# Returns only the selected columns als labels (other columns will not be shown in the facet strips)
first.col.labeller <- function(labels) {

  return(labels[1])

}
```



```{r define_option_templates, echo = FALSE}
# Experiments to derive the best RMarkdown code chunk options for ggplots
# https://stackoverflow.com/questions/42742069/how-to-zoom-or-shrink-the-size-of-ggplot-or-simple-plot-in-rmarkdown-output
# For a list of possible chunk options see:
# https://yihui.name/knitr/options/#chunk_options
num.bars <- max(test.case.groups.pivot.base$maturity.rank, na.rm = TRUE)
# opts_template$set(test1 = list(out.height = 2000, out.width = 1200, fig.height = 8))
opts_template$set(default_opts = list(out.width = 1000,
                                      fig.height = 0.65 * num.bars,   # care for enough height and similar bar widths (e. g. 0.6 * #bars)
                                      fig.align = 'left'))

# How to use the template in a chunk:
# {r, echo = FALSE, opts.label = 'default_opts'}
```





# Disclaimer

This and the accompanying reports show the compliance level of different test set-ups
(mainly a combination of a database system, R DBI driver package and the operating system).

Do not draw conclusions about the stability, maturity, ready-for-production-ness or any type of comparisson
(even between different test set-ups) without in-depth analysis of your your requirements, current situation
and indiviual requirements.

Especially these reports are not meant to blame anybody about anything.



# TODO include DBI intro and use cases from readme file



# How to read the results

## Test case vs. assertions

The DBI conformity tests are done using the R package `DBItest` which uses the package `testthat` internally
to execute the test cases.

To test results in this report are based on a **per test case count**, where test case means:

1. **Test case ("TC")**:      A call of the `testthat::test_that()` function

Internally every test case has many assertions ("asserts")
whereby an assertion is a call of an `testthat::expect*()` function.

The **"per test case count"** gives a good overview which features are **completely** working as expected.



## Semantics of the used colors

There are only three colors used to mark test result rates (percentages):

1. Green: Good or perfekt
2. Orange: OK, but could be better
3. Red: A severe problem
4. Blue: Partially or completely skipped test cases



# Summary of test results

All test results in this report are based on a **per test case count** (if not otherwise stated).

Each test result is based on a test configuration.

The test configurations are shown on the vertical axis.


## Ranked by total DBI-compliance


```{r Total Summary, echo = FALSE, fig.align = 'left', opts.label = 'default_opts'}
TC.result.colors <- eval(formals(derive.strings.HTML.color)$string.colors)   # DIRTY reuse default value of function argument

# To order by the maturity rank (= best results first) the sorted factor is required
data <- test.case.groups.pivot.base[granularity == "All"][order(-maturity.rank)] # reverse sort to show best result on top (coord_flip!)
data[, test.config.label := factor(test.config.label, levels = unique(test.config.label))]

d <- ggplot(data,
  aes(x = test.config.label, y = TC.pct), fill = TC.result) +
  ggtitle("Total result per test configuration (ranked by DBI-compliance)") +
  geom_bar(stat = "identity", aes(fill = TC.result), position = position_stack(), width = 0.7) +
  # geom_histogram(stat = "identity", aes(fill = TC.result), position = position_stack(), width = 0.7, binwidth = 10) + # does not work
  # number of test cases as label
  # Important: Requires data rows in the same order as test.group factors to show the TC.pct in the correct bar!
  #   Done before by sorting the data rows...
  geom_text(aes(x = test.config.label, y = TC.pct, label = test.cases),
            position = position_stack(vjust = 0.5), color = "white", fontface = "bold", size = 3.5) +
  ylab("Test case weights (in %) and counts") +
  # now axis label (redundant to facet name)
  xlab("") + # xlab("Test Case Group") +
  # assign intuitive colors to stacked bars
  scale_fill_manual(values = TC.result.colors, name = "Test case result:") +
  coord_flip() +
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  # scale_y_continuous(labels = scales::percent) +  # percent labels
  # theme(plot.title = element_text(size = 12)) +     # ggtitle font size
  report_theme() +
  # Show total summary and details per test case group in two plots below each other
  # facet_grid( DB.name + DB.version + DBI.driver.pkg ~ ., scales = "free_y", space = "free_y")
  ##### facet_grid(client.OS.name ~ ., scales = "free_y", space = "free_y")  # facet label right
  ######### facet_wrap(~ client.OS.name, ncol = 1, scales = "free_y")   # facet label on top
  facet_grid(client.OS.name.generic ~ ., scales = "free", space = "free", labeller = first.col.labeller)
  # date, DB.name, DB.version, DBI.driver.pkg, client.OS.name
d # show the plot
```



## Ranking per database

Compares different test configurations for the same database to show the best configuration.

```{r Total Summary per DBMS, echo = FALSE, fig.align = 'left', opts.label = 'default_opts'}
TC.result.colors <- eval(formals(derive.strings.HTML.color)$string.colors)   # DIRTY reuse default value of function argument

# To order by the maturity rank (= best results first) the sorted factor is required
data <- test.case.groups.pivot.base[granularity == "All"][order(-maturity.rank)] # reverse sort to show best result on top (coord_flip!)
data[, test.config.label := factor(test.config.label, levels = unique(test.config.label))]

d <- ggplot(data,
  aes(x = test.config.label, y = TC.pct), fill = TC.result) +
  ggtitle("Total result per DBMS ranked by DBI-compliance") +
  geom_bar(stat = "identity", aes(fill = TC.result), position = position_stack(), width = 0.7) +
  # geom_histogram(stat = "identity", aes(fill = TC.result), position = position_stack(), width = 0.7, binwidth = 10) + # does not work
  # number of test cases as label
  # Important: Requires data rows in the same order as test.group factors to show the TC.pct in the correct bar!
  #   Done before by sorting the data rows...
  geom_text(aes(x = test.config.label, y = TC.pct, label = test.cases),
            position = position_stack(vjust = 0.5), color = "white", fontface = "bold", size = 3.5) +
  ylab("Test case weights (in %) and counts") +
  # now axis label (redundant to facet name)
  xlab("") + # xlab("Test Case Group") +
  # assign intuitive colors to stacked bars
  scale_fill_manual(values = TC.result.colors, name = "Test case result:") +
  coord_flip() +
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  # scale_y_continuous(labels = scales::percent) +  # percent labels
  # theme(plot.title = element_text(size = 12)) +     # ggtitle font size
  report_theme() +
  # Show total summary and details per test case group in two plots below each other
  # facet_grid( DB.name + DB.version + DBI.driver.pkg ~ ., scales = "free_y", space = "free_y")
  ##### facet_grid(client.OS.name + DB.name ~ ., scales = "free_y", space = "free_y")
  ######### facet_wrap(~ client.OS.name + DB.name, ncol = 1, scales = "free_y")   # facet label on top
  facet_grid(client.OS.name.generic + DB.name ~ ., scales = "free", space = "free", labeller = first.col.labeller)
  # date, DB.name, DB.version, DBI.driver.pkg, client.OS.name
d # show the plot
```



## Ranking per DBI driver

Compares different test configurations for the same DBI driver to show the best configuration


```{r Total Summary per DBI driver, echo = FALSE, fig.align = 'left', opts.label = 'default_opts'}
# out.width=1000 is perfect

TC.result.colors <- eval(formals(derive.strings.HTML.color)$string.colors)   # DIRTY reuse default value of function argument

# To order by the maturity rank (= best results first) the sorted factor is required
data <- test.case.groups.pivot.base[granularity == "All"][order(-maturity.rank)] # reverse sort to show best result on top (coord_flip!)
data[, test.config.label := factor(test.config.label, levels = unique(test.config.label))]

d <- ggplot(data,
  aes(x = test.config.label, y = TC.pct), fill = TC.result) +
  ggtitle("Total result per DBI driver package ranked by DBI compliance") +
  geom_bar(stat = "identity", aes(fill = TC.result), position = position_stack(), width = 0.75) +
  # geom_histogram(stat = "identity", aes(fill = TC.result), position = position_stack(), width = 0.7, binwidth = 0.01) + # does not work (warning -> ignore binwidth)
  # number of test cases as label
  # Important: Requires data rows in the same order as test.group factors to show the TC.pct in the correct bar!
  #   Done before by sorting the data rows...
  geom_text(aes(x = test.config.label, y = TC.pct, label = test.cases),
            position = position_stack(vjust = 0.5), color = "white", fontface = "bold", size = 3.0) +
  ylab("Test case weights (in %) and counts") +
  # now axis label (redundant to facet name)
  xlab("") + # xlab("Test Case Group") +
  # assign intuitive colors to stacked bars
  scale_fill_manual(values = TC.result.colors, name = "Test case result:") +
  coord_flip() +
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  # scale_y_continuous(labels = scales::percent) +  # percent labels
  report_theme() +
  # theme(plot.title = element_text(size = 10),
  #       axis.text = element_text(size = 8),
  #       strip.text = element_text(size = 8),
  #       axis.title = element_text(size = 8),
  #       legend.text = element_text(size = 8),
  #       legend.title = element_text(size = 10),
  #       legend.position = "bottom") +     # ggtitle font size
  # Show total summary and details per test case group in two plots below each other
  # facet_grid( DB.name + DB.version + DBI.driver.pkg ~ ., scales = "free_y", space = "free_y")
  #### facet_grid(client.OS.name + DBI.driver.pkg ~ ., scales = "free_y", space = "free_y")
  # facet_wrap(~ client.OS.name + DBI.driver.pkg, ncol = 1, scales = "free_y")   # facet label on top
  facet_grid(client.OS.name.generic + DBI.driver.pkg ~ ., scales = "free", space = "free", labeller = first.col.labeller)
  # date, DB.name, DB.version, DBI.driver.pkg, client.OS.name
d # show the plot
```



# Test result reports for each test configuration

```{r, echo = FALSE}
data <- test.case.groups.pivot.base[granularity == "All",
                                    .(report.file.name = make.single.test.HTML.report.file.name(c(.BY, .SD)),
                                      count = .N) ,
                                    by = .(test.config.ID,date, DBI.driver.pkg, DB.name, DB.version, client.OS.name.generic, OS.driver.name, test.config.label)]
# data[, report.file.name := make.single.test.HTML.report.file.name(data)]
data[, report.link := paste0("[Show](", report.file.name, ")")]

data2 <- data[, .("Config.ID" = test.config.ID,
                  "client OS" = client.OS.name.generic,
                  DB.name, DB.version,
                  DBI.driver.pkg,
                  OS.driver.name,
                  report = report.link)
              ][order(Config.ID)]         # [order(`client OS`, DB.name, DB.version, DBI.driver.pkg)]

data2 %>%
kable("html", escape = FALSE)  %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```



# Test results matrix per test case group and test configs

TODO Compare test case groups over all test configs, per DBMS and per DBI driver
