---
title: "DBI compliance test results - comparative report"
author: "https://github.com/aryoda/R_DBI_DB_conformity_reports"
date: '`r format(Sys.Date(), format = "%B %d, %Y")`'
output: 
  html_document: 
    number_sections: yes
    toc: yes
---

<style type="text/css">
div#TOC li {
    list-style:none;
    background-image:none;
    background-repeat:none;
    background-position:0; 
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(knitr)
require(kableExtra)
require(formattable)
require(ggplot2)
```







# How to read the results

## Test case vs. assertions

The DBI conformity tests are done using the R package `DBItest` which uses the package `testthat` internally
to execute the test cases.

To test results in this report are based on a **per test case count**, where test case means:

1. **Test case ("TC")**:      A call of the `testthat::test_that()` function

Internally every test case has many assertions ("asserts")
whereby an assertion is a call of an `testthat::expect*()` function.

The **"per test case count"** gives a good overview which features are **completely** working as expected.



## Semantics of the used colors

There are only three colors used to mark test result rates (percentages):

1. Green: Good or perfekt
2. Orange: OK, but could be better
3. Red: A severe problem
4. Blue: Partially or completely skipped test cases



# Summary of test results

## Ranked by total compliance/maturity


```{r Total Summary, echo = FALSE, fig.align = 'left'}
TC.result.colors <- eval(formals(derive.strings.HTML.color)$string.colors)   # DIRTY reuse default value of function argument

# https://stackoverflow.com/questions/10834382/ggplot2-keep-unused-levels-barplot
# https://stackoverflow.com/questions/30739602/ggplot-reorder-stacked-bar-plot-based-on-values-in-data-frame
# https://stackoverflow.com/questions/37817809/r-ggplot-stacked-bar-chart-with-counts-on-y-axis-but-percentage-as-label
# https://stackoverflow.com/questions/37817809/r-ggplot-stacked-bar-chart-with-counts-on-y-axis-but-percentage-as-label
d <- ggplot(test.case.groups.pivot.base[granularity == "All"],
  aes(x = test.group, y = TC.pct), fill = TC.result) +
  geom_bar(stat = "identity", aes(fill = TC.result), position = position_stack(), width = 0.7) +
  # number of test cases as label
  # Important: Requires data rows in the same order as test.group factors to show the TC.pct in the correct bar!
  #   Done before by sorting the data rows...
  geom_text(aes(x = test.group, y = TC.pct, label = test.cases),
            position = position_stack(vjust = 0.5), color = "white", fontface = "bold", size = 3.5) +
  ylab("Weight in %") +
  # now axis label (redundant to facet name)
  xlab("") + # xlab("Test Case Group") +
  # assign intuitive colors to stacked bars
  scale_fill_manual(values = TC.result.colors, name = "TC result") +
  coord_flip() +
  ggtitle("Total and per test case group (test case weights and counts)") +
  # Show total summary and details per test case group in two plots below each other
  facet_grid( DB.name + DB.version + DBI.driver.pkg ~ ., scales = "free_y", space = "free_y")
  # date, DB.name, DB.version, DBI.driver.pkg, client.OS.name
d # show the plot
```



## Ranking per database

Compares different test configurations for the same database to show the best configuration.


## Ranking per DBI driver

Compares different test configurations for the same DBI driver to show the best configuration
